---
title: "COVID-19 by location"
author: "Danny Hanson"
date: "6/19/2020"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## The Data
All COVID-19 data comes from the [John's Hopkins CSSE repository](https://github.com/CSSEGISandData/COVID-19/). The population data was pulled from the following sources:
- [United State Census](https://www.census.gov/data/datasets/time-series/demo/popest/2010s-counties-total.html)
- [Canada Population](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710000901)
- [Australia Population](https://www.worldatlas.com/articles/the-largest-states-and-territories-of-australia.html)
- [China Population](https://www.worldatlas.com/articles/chinese-provinces-by-population.html)
- [World Population](https://worldpopulationreview.com/)

The script for pulling/joining/cleaning the data can be [found on my github](https://github.com/danhan52/covid19_visuals/blob/master/covid_downloaders.R) and the data itself can also be [pulled from github](https://raw.githubusercontent.com/danhan52/covid19_visuals/master/data/covid.RData).

```{r load_data}
library(shiny)
library(plotly)
library(tidyverse)
library(lubridate)
library(kableExtra)
options(scipen = 100)

load(url('https://raw.githubusercontent.com/danhan52/covid19_visuals/master/data/covid.RData'))

covid <-
  covid_long %>%
  group_by(loc, loc_type, metric) %>%
  mutate(
    delta = people - lag(people, order_by = date, default = 0),
    people_per = round(people / pop * 1000, 2),
    delta_per = round(delta / pop * 1000, 2)
  ) %>%
  ungroup()
```

## Most Recent Data
Below is the most up to date data for the locations shown below.
```{r}
covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    date == max(date)
  ) %>%
  pivot_wider(names_from = metric, values_from = c(people, delta, people_per, delta_per)) %>%
  select(
    Date = date,
    Location = loc,
    Population = pop,
    'Confirmed Cases' = people_confirmed,
    'Deaths' = people_deaths,
    'Increase in Confirmed Cases' = delta_confirmed,
    'Increase in Deaths' = delta_deaths,
    'Confirmed Cases (per 1000 people)' = people_per_confirmed,
    'Deaths (per 1000 people)' = people_per_deaths
  ) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## COVID-19 Cases/Deaths {.tabset .tabset-fade .tabset-pills}
### Confirmed Cases
Below you can see the number of total (cumulative) cases over time.
```{r cases_tot}
p_ct <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'confirmed'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_line(aes(y = people))
ggplotly(p_ct, width = 800, height = 500)
```

Below is the daily change in number of cases with a smoothed curve
```{r cases_chng}
p_cc <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'confirmed'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_point(aes(y = delta), stat = 'identity') +
  geom_smooth(aes(y = delta))
ggplotly(p_cc, width = 800, height = 500)
```

### Deaths
Below you can see the number of total (cumulative) deaths over time.

```{r death_tot}
p_ct <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'deaths'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_line(aes(y = people))
ggplotly(p_ct, width = 800, height = 500)
```

Below is the daily change in number of deaths with a smoothed curve
```{r death_chng}
p_cc <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'deaths'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_point(aes(y = delta), stat = 'identity') +
  geom_smooth(aes(y = delta))
ggplotly(p_cc, width = 800, height = 500)
```

### Confirmed Cases per 1000 people
Below you can see the number of total (cumulative) cases per 1000 people over time.
```{r cases_tot_per}
p_ct <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'confirmed'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_line(aes(y = people_per))
ggplotly(p_ct, width = 800, height = 500)
```

Below is the daily change in number of cases per 1000 people with a smoothed curve
```{r cases_chng_per}
p_cc <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'confirmed'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_point(aes(y = delta_per), stat = 'identity') +
  geom_smooth(aes(y = delta_per))
ggplotly(p_cc, width = 800, height = 500)
```

### Deaths per 1000 people
Below you can see the number of total (cumulative) deaths per 1000 people over time.

```{r death_tot_per}
p_ct <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'confirmed'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_line(aes(y = people_per))
ggplotly(p_ct, width = 800, height = 500)
```

Below is the daily change in number of deaths per 1000 people with a smoothed curve
```{r death_chng_per}
p_cc <-
  covid %>%
  filter(
    loc %in% c('Minnesota', 'Quebec', 'Cyprus', 'Japan', 'Grand Traverse County, Michigan'),
    metric %in% 'confirmed'
  ) %>%
  ggplot(aes(x = date, color = loc)) +
  geom_point(aes(y = delta_per), stat = 'identity') +
  geom_smooth(aes(y = delta_per))
ggplotly(p_cc, width = 800, height = 500)
```


```{r}
covid_us <- 
  c('confirmed' = 'confirmed', 'deaths' = 'deaths') %>%
  map(
    function(x) {
      paste0(
        'https://raw.githubusercontent.com/CSSEGISandData/COVID-19/',
        'master/csse_covid_19_data/csse_covid_19_time_series/',
        'time_series_covid19_',
        x,
        '_US.csv'
      )
    }
  ) %>%
  map_df(function(x) read_csv(url(x)), .id = 'metric') %>%
  pivot_longer(cols = `1/22/20`:last_col(),
               names_to = 'date',
               values_to = 'people') %>%
  mutate(
    fipstr = str_pad(FIPS, 5, 'left', '0'),
    date = as.Date(date, format = '%m/%d/%y')
  )

# https://www.census.gov/data/datasets/time-series/demo/popest/2010s-counties-total.html#par_textimage_70769902
us_pop <-
  read_csv(url(paste0('https://www2.census.gov/programs-surveys/popest/',
                      'datasets/2010-2019/counties/totals/',
                      'co-est2019-alldata.csv'))) %>%
  rename_all(tolower) %>%
  select(stname, ctyname, popestimate2019, state, county) %>%
  mutate(
    ctyname = str_replace(ctyname, ' County', ''),
    fipstr = paste0(state, county)
  )

covid_us <-
  covid_us %>%
  left_join(us_pop, by = 'fipstr') %>%
  group_by(metric, fipstr) %>%
  mutate(delta = people - lag(people, order_by = date, default = 0)) %>%
  ungroup()


url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-states-fips.json'
counties <- rjson::fromJSON(file=url)
df <-
  covid_us %>%
  filter(metric == 'confirmed', date == max(date, na.rm = T), delta >= 0)
fig <- plot_ly(data = df) %>%
  add_trace(
    type="choroplethmapbox",
    geojson=counties,
    locations=~fipstr,
    z=~delta,
    # colorscale="Viridis",
    # zmin=0,
    # zmax=max(df$people, na.rm = T),
    marker=list(line=list(
      width=0),
      opacity=0.5
    )
  ) %>% layout(
    mapbox=list(
      style="carto-positron",
      zoom =2,
      center=list(lon= -95.71, lat=37.09))
  )
# fig

county_df <- map_data("county")
state_df <- map_data("state")

covid_us_plt <-
  covid_us %>%
  filter(metric == 'confirmed', date == max(date, na.rm = T)) %>%
  mutate(join_cty = tolower(Admin2), join_st = tolower(Province_State)) %>%
  inner_join(county_df, by = c('join_cty'='subregion', 'join_st'='region'))

covid_us_plt <-
  covid_us %>%
  filter(metric == 'confirmed', date == max(date, na.rm = T)) %>%
  group_by(Province_State, Country_Region, metric, date) %>%
  summarize(
    people = sum(people, na.rm = T),
    delta = sum(delta, na.rm = T),
    popestimate2019 = sum(popestimate2019, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(join_st = tolower(Province_State)) %>%
  inner_join(state_df, by = c('join_st'='region'))

# create state boundaries
p <- ggplot(covid_us_plt, aes(long, lat, group = group, fill = delta)) +
  geom_polygon(colour = alpha("black", 1/2)) +
  # geom_polygon(data = state_df, colour = "black", fill = NA) + 
  theme_void() +
  scale_fill_viridis_c(direction = -1)

ggplotly(p)

covid_us %>%
  filter(metric == 'confirmed', date == max(date, na.rm = T), Province_State %in% c('Minnesota', 'Michigan')) %>%
  select(Province_State, Admin2, people, delta, everything())

nyt_covid <-
  read_csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')) %>%
  left_join(us_pop, by = c('fips'='fipstr'))

nyt_covid %>%
  filter(`state.x` == 'Michigan') %>%
  distinct(`county.x`) %>%
  arrange(`county.x`)
```

